---
title: "IRCC Permanent Resident Admissions Monthly Timeseries, January 2015 to May 2019"
output: 
  html_notebook: 
    toc: yes
---

# Author

* Jordan Bell
* July 24, 2019
* <https://jordanbell2357.github.io/IRCC.nb.html>

```{r}
library(forecast)
library(pracma)
library(RTransferEntropy)
library(lmtest)
```

Immigration, Refugees and Citizenship Canada (IRCC), [Permanent Residents – Monthly IRCC Updates](https://open.canada.ca/data/en/dataset/f7e5498e-0ad8-4417-85c9-9b8aff9b9eda)

> Canada - Admissions of Permanent Residents by Province/Territory, Census Division and Census Subdivision of the Intended Destination (2018 ranking), January 2015 - May 2019

# Seasonal Decomposition

```{r}
IRCC_admissions <- read.csv("IRCC_admissions.csv")
```

```{r}
Toronto.ts <- ts(IRCC_admissions$Toronto, frequency = 12, start = 2015)

Toronto.ts.components <- decompose(Toronto.ts)

autoplot(Toronto.ts.components)
```

```{r}
ON.ts <- ts(IRCC_admissions$ON, frequency = 12, start = 2015)

ON.ts.components <- decompose(ON.ts)

autoplot(ON.ts.components)
```

```{r}
Vancouver.ts <- ts(IRCC_admissions$Vancouver, frequency = 12, start = 2015)

Vancouver.ts.components <- decompose(Vancouver.ts)

autoplot(Vancouver.ts.components)
```

```{r}
BC.ts <- ts(IRCC_admissions$BC, frequency = 12, start = 2015)

BC.ts.components <- decompose(BC.ts)

autoplot(BC.ts.components)
```


# Cross-Correlation

E. E. Holmes, M. D. Scheuerell, and E. J. Ward, [Applied Time Series Analysis for Fisheries and Environmental Sciences. Section 4.4: Correlation within and among time series](https://nwfsc-timeseries.github.io/atsa-labs/sec-tslab-correlation-within-and-among-time-series.html)

```{r}
length(Toronto.ts)
```

```{r}
ccf(Toronto.ts, Vancouver.ts, main = NA, lag = 52, ylab="Cross-correlation")
```

```{r}
ccf(Toronto.ts, ON.ts, main = NA, lag = 52, ylab="Cross-correlation")
```

```{r}
ccf(Vancouver.ts, BC.ts, main = NA, lag = 52, ylab="Cross-correlation")
```

```{r}
ccf(ON.ts, BC.ts, main = NA, lag = 52, ylab="Cross-correlation")
```



# Sample Entropy

pracma, [Approximate And Sample Entropy](https://www.rdocumentation.org/packages/pracma/versions/1.9.9/topics/approx_entropy)

[Sample Entropy with Rcpp](https://www.r-bloggers.com/sample-entropy-with-rcpp/)

```{r}
sample_entropy(Toronto.ts)

sample_entropy(ON.ts)

sample_entropy(Vancouver.ts)

sample_entropy(BC.ts)

approx_entropy(Toronto.ts)

approx_entropy(ON.ts)

approx_entropy(Vancouver.ts)

approx_entropy(BC.ts)
```

Simon Behrendt, Thomas Dimpfl, Franziska J. Peter, and David J. Zimmermann,
[RTransferEntropy](https://cran.r-project.org/web/packages/RTransferEntropy/vignettes/transfer-entropy.html)

```{r}
calc_te(Toronto.ts, Vancouver.ts)
```

```{r}
calc_te(Toronto.ts, ON.ts)
```

```{r}
calc_te(ON.ts, BC.ts)
```

# Granger Causality

lmtest, [Test For Granger Causality](https://www.rdocumentation.org/packages/lmtest/versions/0.9-37/topics/grangertest)

```{r}
grangertest(Toronto.ts, ON.ts)
```

```{r}
grangertest(Vancouver.ts, BC.ts)
```

```{r}
grangertest(Toronto.ts, Vancouver.ts)
```

```{r}
grangertest(ON.ts, BC.ts)
```
