---
title: "OCHPP Toronto Public Health Datasets"
output: 
  html_notebook: 
    toc: yes
---

# Author

* Jordan Bell
* July 1, 2019
* <https://jordanbell2357.github.io/OCHPP.nb.html>

# Source and descriptions of datasets

**Ontario Community Health Profiles Partnership (OCHPP)**    
<http://www.ontariohealthprofiles.ca/>

**Data — LHIN 7 (Toronto Central and City of Toronto) Neighbourhoods, Ontario Sub-Regions and LHINs**

We use three datasets:

* **Primary Care: Enrolment and Continuity of Care 2011/13**  
1_pc_Continuity_NonRostered_Rostered_Patients_neighb_2013_LHIN_7.xlsx
* **Adult Health and Disease 2016/17**  
1_AHD_2017_RPDB_Neighb_LHIN_7.xlsx
* **Sexual Health: Chlamydia cases, Gonorrhea cases 2013/16**  
2_sh_neighb_chlam_gonor_2013-2016_LHIN_7.xls

We quote from the description of the data:

**The Registered Persons Database (RPDB) as the source of population**  
<http://www.ontariohealthprofiles.ca/o_documents/aboutTheDataON/RPDB_vs_Census_Oct_26_2018.pdf>

> **Two main population sources**
>
>There are two main sources that provide the most reliable estimates of Ontario's population. For
purposes of OCHPP reporting, we use the Registered Persons Database (RPDB) as the source
(denominator) for all of our health-related indicators. While another major source for identifying the
population (denominator) is Statistics Canada’s Census counts (Census), the OCHPP has determined that
the RPDB provides a more consistent measure for the calculation of rates for health indicators and
conditions. We explain why and provide examples to illustrate.
>
> **RPDB**
>
>The RPDB (database) provides basic demographic information about anyone who has ever received an
Ontario Health Insurance Plan (OHIP) card. OHIP cards include a unique Health Card Number (HCN) to
identify a person's age, sex and address, including postal code. The postal codes used at ICES come
mainly from HCNs. Information from the health card is stored in the RPDB. Health cards are usually
renewed every 5 years, a process that helps to ensure that information in the system is periodically
refreshed, such as address, for example, should a person move and update the card within that time
frame. The system may also be updated more frequently if an individual interacts with the healthcare
system between renewal periods. This allows for a more current source of location-based data for
health reporting.
>
> **Census**
>
>Another source of data on where people live can come from Statistics Canada (Stats Can). Stats Can
collects data on individuals living in Canada through the Census, a survey conducted every 5 years. Stats Can
reports who lives in a given area at one point in time – i.e. based on the information an individual provides when
completing the Census
>
> **RPDB vs. Census**
>
>While at the provincial level, the number of people we identify in the RPDB as living in Ontario does not
differ much from Census estimates, differences are more pronounced at smaller areas such as
neighbourhoods or local areas. This particular issue is most evident in areas of high migration where we
have observed large differences in rates using RPDB vs. the Census in part due to population mobility.  
>For example, in some areas of Ontario, particularly in larger urban centres, newcomers to the area often
settle first “downtown” but over time, may move from downtown to outlying areas. The majority of
people who move do not change or update their health card until renewal time so they stay in the RPDB
with their original health card information including the address and postal code of their “downtown”
address.  
>The same is true of Census population counts: when someone moves, they are still considered living at
their original address and comprise the population of that area until the next Census is taken (every 5
years).

# Shapefile for City of Toronto's 140 neighbourhoods

We now import and process a shapefile for Toronto's 140 neighbourhoods.

<https://www.toronto.ca/city-government/data-research-maps/open-data/open-data-catalogue/#a45bd45a-ede8-730e-1abc-93105b2c439f>

<http://opendata.toronto.ca/gcc/neighbourhoods_planning_areas_wgs84.zip>

**Neighbourhoods**

> Owner: Social Development, Finance & Administration
>
> Currency: June 2014

**Neighbourhoods (WGS84)**

```{r}
library(data.table) #fread
library(plyr) #join
library(ggplot2) #ggplot, fortify
library(sp) #used by rgdal
library(rgdal) #readOGR
library(rgeos) #gCentroid
library(scales) #scale_fill_distiller, percent_format
library(ggmap) #theme_nothing
library(Hmisc) #rcorr
```
Input shapefile
```{r}
nbds.sh <- readOGR("C:/Users/14165/Desktop/ArcGIS/SHAPEFILES/neighbourhoods_planning_areas_wgs84", "NEIGHBORHOODS_WGS84")
```
Add "id" column
```{r}
nbds.sh@data$id <- as.integer(nbds.sh@data$AREA_S_CD)
```
Make centroids of each neighbourhood, for placing labels when plotting
```{r}
nbds.sh.centroids  <- as.data.frame(gCentroid(nbds.sh, byid = TRUE))
```
Add "id" column
```{r}
nbds.sh.centroids$id <- nbds.sh@data$id
```
Shapefile processing
```{r}
nbds.sh.points = fortify(nbds.sh, region = "id")

nbds.sh.df = join(nbds.sh.points, nbds.sh@data, by = "id")
```
We are now done with shapefile processing.

We copied selected columns from sheets of the .xls/.xlsx files to new sheets, renamed columns by changing spaces and hyphens to underscores and spelling symbols, and exported these to .csv files.

We shall visualize selected data from these tables on a map. Then we compute correlations.

# Primary Care: Enrolment and Continuity of Care 2011/13

1_pc_Continuity_NonRostered_Rostered_Patients_neighb_2013_LHIN_7.xlsx!Continuity_Enrolled_NonEnrolled:

> Primary Care: Enrolment and Continuity of Care (Both sexes, Ages 19+) for Toronto Neighbourhoods and Toronto Central LHIN, 2011/12 to 2012/13 (April 1, 2011-March 31, 2013)
>
> Data sources: Population - Ontario Heath Insurance Plan (OHIP) physician claims, Client Agency Provider Enrolment (CAPE) tables, Registered Persons' Database, Community Health Centre client encounter data.
>
> Population Enrolled: rostered with a Patient Enrolment Model (PEM) or registered with a Community Health Centre (CHC).	
>
> Population Non-Enrolled: not rostered with a Patient Enrolment Model (PEM) or registered with a Community Health Centre (CHC).			

"id", "Population_19_plus", "Enrolled_Population", "Enrolled_Population_with_No_Visits", "Enrolled_Population_with_1_or_2_Visits", "Enrolled_Population_with_3_plus_Visits", "Non_Enrolled_Population", "Non_Enrolled_Population_with_No_Visits", "Non_Enrolled_Population_with_1_or_2_Visits", "Non_Enrolled_Population_with_3_plus_Visits", "Total_Population_Enrolled_and_Non_Enrolled", "Total_Population_with_No_Visits", "Total_Population_with_1_or_2_Visits", "Total_Population_with_3_plus_Visits"

**1_pc_Continuity_NonRostered_Rostered_Patients_neighb_2013_LHIN_7.csv**

```{r}
patients <- fread("1_pc_Continuity_NonRostered_Rostered_Patients_neighb_2013_LHIN_7.csv")
```

```{r}
str(patients)
```


Create ratio columns for selected variables
```{r}
patients$Ratio_Non_Enrolled_Population <- patients$Non_Enrolled_Population/patients$Population_19_plus

patients$Ratio_Total_Population_with_No_Visits <- patients$Total_Population_with_No_Visits/patients$Population_19_plus

patients$Ratio_Total_Population_with_3_plus_Visits <- patients$Total_Population_with_3_plus_Visits/patients$Population_19_plus

patients_ratios <- patients[,.(id,Ratio_Non_Enrolled_Population,Ratio_Total_Population_with_No_Visits,Ratio_Total_Population_with_3_plus_Visits)]
```

## Ratio_Non_Enrolled_Population
```{r}
Ratio_Non_Enrolled_Population <- patients_ratios[, .(id,Ratio_Non_Enrolled_Population)]

Ratio_Non_Enrolled_Population.sh <- merge(nbds.sh.df, Ratio_Non_Enrolled_Population, by = "id")
```
Make graphics object
```{r}
p.Ratio_Non_Enrolled_Population <- ggplot() +
  geom_polygon(data = Ratio_Non_Enrolled_Population.sh, 
               aes(x = long, y = lat, group = group, fill = Ratio_Non_Enrolled_Population), 
               color = "black", size = 0.2) + 
  coord_map() + 
  scale_fill_distiller(name="Ratio", labels=percent_format(accuracy=1), palette = "PuRd", trans = "reverse", breaks = pretty_breaks(n = 8))+
  theme_nothing(legend = TRUE) + 
  labs(title="Ratio of total population non-enrolled") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = nbds.sh.centroids, size = 2)
```
Plot graphics object
```{r}
p.Ratio_Non_Enrolled_Population + guides(fill = guide_legend(reverse=TRUE))
```

## Ratio_Total_Population_with_No_Visits
```{r}
Ratio_Total_Population_with_No_Visits <- patients_ratios[, .(id,Ratio_Total_Population_with_No_Visits)]

Ratio_Total_Population_with_No_Visits.sh <- merge(nbds.sh.df, Ratio_Total_Population_with_No_Visits, by = "id")
```

Make graphics object
```{r}
p.Ratio_Total_Population_with_No_Visits <- ggplot() +
  geom_polygon(data = Ratio_Total_Population_with_No_Visits.sh, 
               aes(x = long, y = lat, group = group, fill = Ratio_Total_Population_with_No_Visits), 
               color = "black", size = 0.2) + 
  coord_map() + 
  scale_fill_distiller(name="Ratio", labels=percent_format(accuracy=1), palette = "PuRd", trans = "reverse", breaks = pretty_breaks(n = 8)) + 
  theme_nothing(legend = TRUE) + 
  labs(title="Ratio of total population with no visits") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = nbds.sh.centroids, size = 2)
```
Plot graphics object
```{r}
p.Ratio_Total_Population_with_No_Visits + guides(fill = guide_legend(reverse=TRUE))
```

## Ratio_Total_Population_with_3_plus_Visits
```{r}
Ratio_Total_Population_with_3_plus_Visits <- patients_ratios[, .(id,Ratio_Total_Population_with_3_plus_Visits)]

Ratio_Total_Population_with_3_plus_Visits.sh <- merge(nbds.sh.df, Ratio_Total_Population_with_3_plus_Visits, by = "id")
```
Make graphics object
```{r}
p.Ratio_Total_Population_with_3_plus_Visits <- ggplot() +
  geom_polygon(data = Ratio_Total_Population_with_3_plus_Visits.sh, 
               aes(x = long, y = lat, group = group, fill = Ratio_Total_Population_with_3_plus_Visits), 
               color = "black", size = 0.2) + 
  coord_map() + 
  scale_fill_distiller(name="Ratio", labels=percent_format(accuracy=1), palette = "Greens", trans = "reverse", breaks = pretty_breaks(n = 8))+
  theme_nothing(legend = TRUE) + 
  labs(title="Ratio of total population with 3+ visits") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = nbds.sh.centroids, size = 2)
```
Plot graphics object
```{r}
p.Ratio_Total_Population_with_3_plus_Visits + guides(fill = guide_legend(reverse=TRUE))
```

# Adult Health and Disease 2016/17

1_AHD_2017_RPDB_Neighb_LHIN_7.xlsx!Diabetes_Neighb_Toronto:

> Demographics - Denominator: Ontario Ministry of Health and Long-Term Care Registered Persons Database (RPDB), population aged 20+ who were alive and living in the Ontario on April 1st, 2016 
>
> Numerator: derived from validated, disease registries maintained by the Institute for Clinical Evaluative Sciences (ICES)

1_AHD_2017_RPDB_Neighb_LHIN_7.xlsx!COPD_Neighb_Toronto:

> Demographics - Denominator: Ontario Ministry of Health and Long-Term Care Registered Persons Database (RPDB), population aged 35+ who were alive and living in the Ontario on April 1st, 2016 
>
> Numerator: derived from validated, disease registries maintained by the Institute for Clinical Evaluative Sciences (ICES)

"id", "People_with_Diabetes_20_plus", "Total_Population_2016_RPDB_20_plus", "Prevalence_per_hundred_Diabetes_20_plus", "People_with_COPD_35_plus", "Total_Population_2016_RPDB_35_plus", "Prevalence_per_hundred_COPD_35_plus"

**1_AHD_2017_RPDB_Neighb_LHIN_7.csv**

```{r}
chronic <- fread("1_AHD_2017_RPDB_Neighb_LHIN_7.csv")

chronic$Ratio_People_with_Diabetes_20_plus <- chronic$People_with_Diabetes_20_plus/chronic$Total_Population_2016_RPDB_20_plus

chronic$Ratio_People_with_COPD_35_plus <- chronic$People_with_COPD_35_plus/chronic$Total_Population_2016_RPDB_35_plus

chronic_ratios <- chronic[,.(id,Ratio_People_with_Diabetes_20_plus,Ratio_People_with_COPD_35_plus)]
```

```{r}
Ratio_People_with_Diabetes_20_plus <- chronic_ratios[, .(id,Ratio_People_with_Diabetes_20_plus)]

Ratio_People_with_Diabetes_20_plus.sh <- merge(nbds.sh.df, Ratio_People_with_Diabetes_20_plus, by = "id")
```

Make and plot graphics object
```{r}
p.Ratio_People_with_Diabetes_20_plus <- ggplot() +
  geom_polygon(data = Ratio_People_with_Diabetes_20_plus.sh, 
               aes(x = long, y = lat, group = group, fill = Ratio_People_with_Diabetes_20_plus), 
               color = "black", size = 0.2) + 
  coord_map() + 
  scale_fill_distiller(name="Ratio", labels=percent_format(accuracy=1), palette = "PuRd", trans = "reverse",
                       breaks = pretty_breaks(n = 8)) + 
  theme_nothing(legend = TRUE)+
  labs(title="Ratio of people with diabetes, 20+") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = nbds.sh.centroids, size = 2)

p.Ratio_People_with_Diabetes_20_plus + guides(fill = guide_legend(reverse=TRUE))
```

```{r}
Ratio_People_with_COPD_35_plus <- chronic_ratios[, .(id,Ratio_People_with_COPD_35_plus)]

Ratio_People_with_COPD_35_plus.sh <- merge(nbds.sh.df, Ratio_People_with_COPD_35_plus, by = "id")
```
Make and plot graphics object
```{r}
p.Ratio_People_with_COPD_35_plus <- ggplot() +
  geom_polygon(data = Ratio_People_with_COPD_35_plus.sh, 
               aes(x = long, y = lat, group = group, fill = Ratio_People_with_COPD_35_plus), 
               color = "black", size = 0.2) + 
  coord_map() + 
  scale_fill_distiller(name="Ratio", labels=percent_format(accuracy=1), palette = "PuRd", trans = "reverse",
                       breaks = pretty_breaks(n = 8))+
  theme_nothing(legend = TRUE) + 
  labs(title="Ratio of people with COPD, 35+") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = nbds.sh.centroids, size = 2)

p.Ratio_People_with_COPD_35_plus + guides(fill = guide_legend(reverse=TRUE))
```

# Sexual Health: Chlamydia cases, Gonorrhea cases 2013/16

2_sh_neighb_chlam_gonor_2013-2016_LHIN_7.xlsx!Chlamydia_Male_Female_Tor_neigh:

>  Number of Chlamydia Cases for Age 15 years and over by Gender (Male, Female), Toronto Neighbourhoods, 2013 to 2016 Calendar Years Combined
>
>  Demographics - Denominator: Based on 2011 Census population estimates. (Statistics Canada, 2011 Census of Population).
>
>  Numerator: Number of Chlamydia cases for 4 year (2013 to 2016) observation period.        Data source: Data as of July 20, 2017, Toronto Public Health integrated Public Health Information System [iPHIS]. 
>
> * Average annual rate of Chlamydia cases (/100,000) Males (2013 to 2016), All Ages 15+
>
> * Average annual rate of Chlamydia cases (/100,000) Females (2013 to 2016), All Ages 15+

"id", "Chlamydia_cases_male", "Population_male", "Chlamydia_cases_female", "Population_female"

**2_sh_neighb_chlam_gonor_2013-2016_LHIN_7.csv**

```{r}
chlamydia <- fread("2_sh_neighb_chlam_gonor_2013-2016_LHIN_7.csv")

chlamydia$Ratio_Chlamydia_cases_male <- chlamydia$Chlamydia_cases_male/chlamydia$Population_male

chlamydia$Ratio_Chlamydia_cases_female <- chlamydia$Chlamydia_cases_female/chlamydia$Population_female

chlamydia_ratios <- chlamydia[,.(id,Ratio_Chlamydia_cases_male,Ratio_Chlamydia_cases_female)]
```

```{r}
Ratio_Chlamydia_cases_male <- chlamydia_ratios[, .(id,Ratio_Chlamydia_cases_male)]

Ratio_Chlamydia_cases_male.sh <- merge(nbds.sh.df, Ratio_Chlamydia_cases_male, by = "id")
```
Make and plot graphics object
```{r}
p.Ratio_Chlamydia_cases_male <- ggplot() +
  geom_polygon(data = Ratio_Chlamydia_cases_male.sh, 
               aes(x = long, y = lat, group = group, fill = Ratio_Chlamydia_cases_male), 
               color = "black", size = 0.2) + 
  coord_map() + 
  scale_fill_distiller(name="Cases/population", labels=percent_format(accuracy=1), palette = "YlOrBr", trans = "reverse", breaks = pretty_breaks(n = 8)) + 
  theme_nothing(legend = TRUE) + 
  labs(title="Number of male chlamydia cases/male population") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = nbds.sh.centroids, size = 2)

p.Ratio_Chlamydia_cases_male + guides(fill = guide_legend(reverse=TRUE))
```

# Correlations
```{r}
OCHPP_ratios <- merge(merge(patients_ratios, chronic_ratios), chlamydia_ratios)
```
Structure of data table:
```{r}
str(OCHPP_ratios)
```
Drop "id" column
```{r}
OCHPP_ratios_noid <- OCHPP_ratios[,-c("id")]

str(OCHPP_ratios_noid)
```

Compute correlations
```{r}
rcorr(as.matrix(OCHPP_ratios_noid), type="pearson")
```

## Linear regression
```{r}
lm_chlam <- lm(Ratio_Chlamydia_cases_male ~ Ratio_Chlamydia_cases_female, data = OCHPP_ratios_noid)
summary(lm_chlam)
```


```{r}
plot(lm_chlam)
```

## Choropleth of residuals 

```{r}
str(lm_chlam$residuals)
```
Make dataframe of residuals with "id" column
```{r}
lm_chlam_df <- data.frame(id = c(1:140), residual = lm_chlam$residuals)
head(lm_chlam_df)
```


```{r}
lm_chlam_df.sh <- merge(nbds.sh.df, lm_chlam_df, by = "id")
```
Make and plot graphics object
```{r}
p.lm_chlam_df <- ggplot() +
  geom_polygon(data = lm_chlam_df.sh, 
               aes(x = long, y = lat, group = group, fill = residual), 
               color = "black", size = 0.2) + 
  coord_map() + 
  scale_fill_distiller(name="Residuals", palette = "Blues", trans = "reverse", breaks = pretty_breaks(n = 8)) + 
  theme_nothing(legend = TRUE) + 
  labs(title="Residuals of male chlamydia cases ~ female chlamydia cases") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = nbds.sh.centroids, size = 2)

p.lm_chlam_df + guides(fill = guide_legend(reverse=TRUE))
```