---
title: "Toronto Water Consumption, 2000-2015"
output: 
  html_notebook: 
    toc: yes
---

```{r}
library(data.table) #fread
library(plyr) #join
library(ggplot2) #ggplot, fortify
library(sp) #used by rgdal
library(rgdal) #readOGR
library(rgeos) #gCentroid
library(scales) #scale_fill_distiller, percent_format
library(ggmap) #theme_nothing
library(forecast) #autoplot ts
library(viridisLite)
library(viridis)
```

# Author

* Jordan Bell
* July 10, 2019
* <https://jordanbell2357.github.io/Toronto_Water.nb.html>

# City of Toronto Water Billing by Ward

[Water Billing by Ward](https://www.toronto.ca/city-government/data-research-maps/open-data/open-data-catalogue/#e2b419c3-a8a8-7949-d80d-02585f4832f5)

CoT_Water_Consumption.ZIP

Water_Consumption_2000.xls - Water_Consumption_2015.xls

Volumes in cubic meters.

"city ward", "year", "residential accounts", "annual residential usage",
"average residential usage", "commercial accounts", "annual commercial usage",
"average commercial usage", "total count", "total consumption", "average consumption"

```{r}
water_2000 <- fread("Water_Consumption_2000.csv")
water_2001 <- fread("Water_Consumption_2001.csv")
water_2002 <- fread("Water_Consumption_2002.csv")
water_2003 <- fread("Water_Consumption_2003.csv")
water_2004 <- fread("Water_Consumption_2004.csv")
water_2005 <- fread("Water_Consumption_2005.csv")
water_2006 <- fread("Water_Consumption_2006.csv")
water_2007 <- fread("Water_Consumption_2007.csv")
water_2008 <- fread("Water_Consumption_2008.csv")
water_2009 <- fread("Water_Consumption_2009.csv")
water_2010 <- fread("Water_Consumption_2010.csv")
water_2011 <- fread("Water_Consumption_2011.csv")
water_2012 <- fread("Water_Consumption_2012.csv")
water_2013 <- fread("Water_Consumption_2013.csv")
water_2014 <- fread("Water_Consumption_2014.csv")
water_2015 <- fread("Water_Consumption_2015.csv")
```

```{r}
v <- c("id", "year", "residential_accounts", "annual_residential_usage",
"average_residential_usage", "commercial_accounts", "annual_commercial_usage",
"average_commercial_usage", "total_count", "total_consumption", "average_consumption")

setnames(water_2000, v)
setnames(water_2001, v)
setnames(water_2002, v)
setnames(water_2003, v)
setnames(water_2004, v)
setnames(water_2005, v)
setnames(water_2006, v)
setnames(water_2007, v)
setnames(water_2008, v)
setnames(water_2009, v)
setnames(water_2010, v)
setnames(water_2011, v)
setnames(water_2012, v)
setnames(water_2013, v)
setnames(water_2014, v)
setnames(water_2015, v)
```

```{r}
head(water_2015)
```

## Shapefile

```{r}
wards.sh <- readOGR("C:/Users/14165/Desktop/Shapefiles/May2010_WGS84", "icitw_wgs84")
```

Add "id" column
```{r}
wards.sh@data$id <- as.integer(wards.sh@data$SCODE_NAME)
```

Make centroids of each ward, for placing labels when plotting
```{r}
wards.centroids  <- as.data.frame(gCentroid(wards.sh, byid = TRUE))
```

Add "id" column
```{r}
wards.centroids$id <- wards.sh@data$id
```

Shapefile processing
```{r}
wards.points = fortify(wards.sh, region="id")

wards.df = join(wards.points, wards.sh@data, by="id")
```


## Choropleths

```{r}
wards.water_2015 <- merge(wards.df, water_2015, by = "id")

p.2015_residential <- ggplot() + 
  geom_polygon(data = wards.water_2015,
               aes(x = long, y = lat, group = group, fill = average_residential_usage),
               color = "black", size = 0.25) +  
  coord_map() + 
  scale_fill_distiller(name="Cubic meters/account", palette = "Blues", trans = "reverse", breaks = pretty_breaks(n = 8)) + 
  theme_nothing(legend = TRUE) + 
  labs(title="Average annual residential water use per account by ward, 2015") + 
  geom_text(aes(x=x, y=y, group=NULL, label=id), data = wards.centroids, size = 2)

p.2015_residential + guides(fill = guide_legend(reverse = TRUE))
```

```{r}
p.2015_commercial <- ggplot() + 
  geom_polygon(data = wards.water_2015,
               aes(x = long, y = lat, group = group, fill = average_commercial_usage),
               color = "black", size = 0.25) +  
  coord_map() + 
  scale_fill_distiller(name="Cubic meters/account", palette = "Blues", trans = "reverse", breaks = pretty_breaks(n = 8)) + 
  theme_nothing(legend = TRUE) + 
  labs(title="Average annual commercial water use per account by ward, 2015") + 
  geom_text(aes(x=x, y=y, group=NULL, label=id), data = wards.centroids, size = 2)

p.2015_commercial + guides(fill = guide_legend(reverse = TRUE))
```

```{r}
water_2015$average_residential_usage_change <- (water_2015$average_residential_usage-water_2000$average_residential_usage)/water_2000$average_residential_usage

water_2015$average_commercial_usage_change <- (water_2015$average_commercial_usage-water_2000$average_commercial_usage)/water_2000$average_commercial_usage
```

```{r}
wards.water_2015 <- merge(wards.df, water_2015, by = "id")

p.residential_change <- ggplot() + 
  geom_polygon(data = wards.water_2015,
               aes(x = long, y = lat, group = group, fill = average_residential_usage_change),
               color = "black", size = 0.25) +  
  coord_map() + 
  scale_fill_viridis(option="magma", labels=percent, name="Change") + 
  theme_nothing(legend = TRUE) + 
  labs(title="Change in Average Annual Residential Water Usage, 2000-2015") + 
  geom_text(aes(x=x, y=y, group=NULL, label=id), data = wards.centroids, size = 2)

p.residential_change + guides(fill = guide_legend(reverse = TRUE))
```

```{r}
p.commercial_change <- ggplot() + 
  geom_polygon(data = wards.water_2015,
               aes(x = long, y = lat, group = group, fill = average_commercial_usage_change),
               color = "black", size = 0.25) +  
  coord_map() + 
  scale_fill_viridis(option="magma", labels=percent, name="Change") + 
  theme_nothing(legend = TRUE) + 
  labs(title="Change in Average Annual Commercial Water Usage, 2000-2015") + 
  geom_text(aes(x=x, y=y, group=NULL, label=id), data = wards.centroids, size = 2)

p.commercial_change + guides(fill = guide_legend(reverse = TRUE))
```


## Time Series

```{r}
commercial_accounts_2000 <- sum(water_2000$commercial_accounts)
  
commercial_usage_2000 <- sum(water_2000$annual_commercial_usage)

commercial_accounts_2001 <- sum(water_2001$commercial_accounts)
  
commercial_usage_2001 <- sum(water_2001$annual_commercial_usage)

commercial_accounts_2002 <- sum(water_2002$commercial_accounts)
  
commercial_usage_2002 <- sum(water_2002$annual_commercial_usage)

commercial_accounts_2003 <- sum(water_2003$commercial_accounts)
  
commercial_usage_2003 <- sum(water_2003$annual_commercial_usage)

commercial_accounts_2004 <- sum(water_2005$commercial_accounts)
  
commercial_usage_2004 <- sum(water_2004$annual_commercial_usage)

commercial_accounts_2005 <- sum(water_2005$commercial_accounts)
  
commercial_usage_2005 <- sum(water_2005$annual_commercial_usage)

commercial_accounts_2006 <- sum(water_2006$commercial_accounts)
  
commercial_usage_2006 <- sum(water_2007$annual_commercial_usage)

commercial_accounts_2007 <- sum(water_2007$commercial_accounts)
  
commercial_usage_2007 <- sum(water_2007$annual_commercial_usage)

commercial_accounts_2008 <- sum(water_2008$commercial_accounts)
  
commercial_usage_2008 <- sum(water_2008$annual_commercial_usage)

commercial_accounts_2009 <- sum(water_2009$commercial_accounts)
  
commercial_usage_2009 <- sum(water_2009$annual_commercial_usage)

commercial_accounts_2010 <- sum(water_2010$commercial_accounts)
  
commercial_usage_2010 <- sum(water_2010$annual_commercial_usage)

commercial_accounts_2011 <- sum(water_2011$commercial_accounts)
  
commercial_usage_2011 <- sum(water_2011$annual_commercial_usage)

commercial_accounts_2012 <- sum(water_2012$commercial_accounts)
  
commercial_usage_2012 <- sum(water_2012$annual_commercial_usage)

commercial_accounts_2013 <- sum(water_2013$commercial_accounts)
  
commercial_usage_2013 <- sum(water_2013$annual_commercial_usage)

commercial_accounts_2014 <- sum(water_2014$commercial_accounts)
  
commercial_usage_2014 <- sum(water_2014$annual_commercial_usage)

commercial_accounts_2015 <- sum(water_2015$commercial_accounts)
  
commercial_usage_2015 <- sum(water_2015$annual_commercial_usage)
```

```{r}
commercial_accounts <- c(commercial_accounts_2000, commercial_accounts_2001, commercial_accounts_2002,
                         commercial_accounts_2003, commercial_accounts_2004, commercial_accounts_2005,
                         commercial_accounts_2006, commercial_accounts_2007, commercial_accounts_2008,
                         commercial_accounts_2009, commercial_accounts_2010, commercial_accounts_2011,
                         commercial_accounts_2012, commercial_accounts_2013, commercial_accounts_2014,
                         commercial_accounts_2015)

commercial_usage <- c(commercial_usage_2000, commercial_usage_2001, commercial_usage_2002,
                      commercial_usage_2003, commercial_usage_2004, commercial_usage_2005,
                      commercial_usage_2006, commercial_usage_2007, commercial_usage_2008,
                      commercial_usage_2009, commercial_usage_2010, commercial_usage_2011,
                      commercial_usage_2012, commercial_usage_2013, commercial_usage_2014,
                      commercial_usage_2015)
```

```{r}
average_commercial_usage <- commercial_usage/commercial_accounts
```

```{r}
commercial_usage.ts <- ts(commercial_usage, start=2000, frequency=1)
```

```{r}
plot(commercial_usage.ts, main = "Total Annual Commercial Usage, cubic meters")
```

```{r}
average_commercial_usage.ts <- ts(average_commercial_usage, start=2000, frequency=1)
```

```{r}
plot(average_commercial_usage.ts, main = "Average Annual Commercial Usage, cubic meters")
```


# Biennial Drinking Water Plants Survey (DKWP)

[Potable water use by sector and average daily use](https://www150.statcan.gc.ca/t1/tbl1/en/cv!recreate.action?pid=3810027101&selectedNodeIds=1D7&checkedLevels=1D1,1D2&refPeriods=20110101,20170101&dimensionLayouts=layout3,layout3,layout2&vectorDisplay=false)

Table: 38-10-0271-01 (formerly CANSIM 153-0127)

3810027101_databaseLoadingData.csv

```{r}
DKWP_Ontario_use <- fread("3810027101_databaseLoadingData.csv")

str(DKWP_Ontario_use)
```

```{r}
unique(DKWP_Ontario_use$`Characteristics of potable water use`)
```

```{r}
DKWP_Ontario_average_daily_per_capita <- DKWP_Ontario_use[`Characteristics of potable water use`=="Average daily total use per capita of the population served", .(Year = REF_DATE, Average_daily_per_capita_liters = VALUE)]

DKWP_Ontario_average_residential <- DKWP_Ontario_use[`Characteristics of potable water use`=="Average daily residential use per capita of the population served", .(Year = REF_DATE, Average_daily_residential_liters = VALUE)]
```

```{r}
DKWP_Ontario_average_daily_per_capita
```

```{r}
DKWP_Ontario_average_residential
```

[Potable water volumes processed by drinking water plants, by month (x 1,000,000)](https://www150.statcan.gc.ca/t1/tbl1/en/cv!recreate.action?pid=3810027201&selectedNodeIds=1D7&checkedLevels=1D2&refPeriods=20050101,20170101&dimensionLayouts=layout3,layout3,layout2&vectorDisplay=false)

Table: 38-10-0272-01 (formerly CANSIM 153-0124)

3810027201_databaseLoadingData.csv

```{r}
DKWP_Ontario_processed <- fread("3810027201_databaseLoadingData.csv")
```

VALUE = millions cubic meters

```{r}
str(DKWP_Ontario_processed)
```

```{r}
DKWP_Ontario_processed$Month <- match(DKWP_Ontario_processed$Month, month.name)

str(DKWP_Ontario_processed)
```

```{r}
DKWP_Ontario_processed_volumes <- DKWP_Ontario_processed[,.(Volume = VALUE), by = .(Year = REF_DATE, Month)]

DKWP_Ontario_processed_volumes <- DKWP_Ontario_processed_volumes[order(Year,Month)]

str(DKWP_Ontario_processed_volumes)
```

```{r}
DKWP_Ontario_processed_volumes.ts <- ts(DKWP_Ontario_processed_volumes$Volume, start = 2005, frequency = 12)
```

```{r}
DKWP_Ontario_processed_volumes.ts.components <- decompose(DKWP_Ontario_processed_volumes.ts)

autoplot(DKWP_Ontario_processed_volumes.ts.components)
```

[Wastewater volumes processed by municipal sewage systems (x 1,000,000)](https://www150.statcan.gc.ca/t1/tbl1/en/cv!recreate.action?pid=3810009901&selectedNodeIds=1D7,2D2,2D3,2D4,2D5,2D6,2D7,2D8,2D9,2D10,2D11,2D12,2D13&checkedLevels=&refPeriods=20130101,20170101&dimensionLayouts=layout3,layout3,layout2&vectorDisplay=false)

Table: 38-10-0099-01

3810009901_databaseLoadingData.csv

VALUE = millions cubic meters

```{r}
DKWP_Ontario_wastewater <- fread("3810009901_databaseLoadingData.csv")

str(DKWP_Ontario_wastewater)
```

```{r}
DKWP_Ontario_wastewater$Month <- match(DKWP_Ontario_wastewater$Month, month.name)

DKWP_Ontario_wastewater_volumes <- DKWP_Ontario_wastewater[,.(Volume = VALUE), by = .(Year = REF_DATE, Month)]

DKWP_Ontario_wastewater_volumes <- DKWP_Ontario_wastewater_volumes[order(Year,Month)]

DKWP_Ontario_wastewater_volumes.ts <- ts(DKWP_Ontario_wastewater_volumes$Volume, start = 2013, frequency = 12)

DKWP_Ontario_wastewater_volumes.ts.components <- decompose(DKWP_Ontario_wastewater_volumes.ts)

autoplot(DKWP_Ontario_wastewater_volumes.ts.components)
```

# Further resources

## Ontario Municipal Wastewater

[Municipal treated wastewater effluent](https://www.ontario.ca/data/municipal-treated-wastewater-effluent)

> Data contains flows and effluent quality from our municipal wastewater treatment plants.

MWD2008.xlsx - MWD2016.xlsx

## City of Toronto, 2019

[Water Consumption Report](https://www.toronto.ca/services-payments/water-environment/tap-water-in-toronto/daily-water-consumption-report/)

## MPMP Multi-Year Report

[MPMP Multi-Year Report - By Municipality](https://efis.fma.csc.gov.on.ca/fir/MPMP%20Multi%20Year.htm)

Toronto C 26-Jan-17: MPMP-MULTI-YR-Toronto C.xls

e.g.

> SLC 91 3112 45	b)	Total costs for the treatment and disposal of wastewater per megalitre.

> SLC 91 3311  45	b)	Total costs for the treatment of drinking water per megalitre.

## Statistics Canada

[2011 Municipal Water Use Report – Municipal Water Use 2009 Statistics](http://www.ec.gc.ca/doc/publications/eau-water/com1454/index-eng.htm), Table 2: Total and Residential per Capita Water Use, by Province/Territory and Municipal Population