---
title: "Toronto Uber Movement and Transportation Tomorrow Survey (TTS) Datasets"
output: 
  html_notebook: 
    toc: yes
---

# Author

* Jordan Bell
* July 24, 2019
* <https://jordanbell2357.github.io/TTS.nb.html>

```{r}
library(data.table)
library(dplyr)
library(sp)
library(sf)
library(ggplot2)
library(viridisLite)
library(viridis)
```

# Uber Movement

[Uber Movement](https://movement.uber.com/?lang=en-CA)

GeoJSON file

toronto_neighbourhoods.json

Dataset for Q2 of 2018

toronto-neighbourhoods-2018-2-All-MonthlyAggregate.csv

```{r}
uber_nbds <- st_read("./toronto_neighbourhoods.json", stringsAsFactors=FALSE)

str(uber_nbds)
```

```{r}
uber_nbds$ID <- as.integer(uber_nbds$MOVEMENT_ID)
```

```{r}
uber_nbds$DISPLAY_NAME
```


## Bay Street Corridor

Bay Street Corridor is City of Toronto Neighbourhood 76. This is 
Uber Neighbourhood 113. We hold fixed this neighbourhood and examine travel times whose destination
is to this neighbourhood.

```{r}
uber_movement <- fread("toronto-neighbourhoods-2018-2-All-MonthlyAggregate.csv")

str(uber_movement)
```

[Introduction to data.table](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html)

```{r}
uber_source_113 <- uber_movement[dstid == 113, .(MEAN_TRIP_DURATION = mean(mean_travel_time)), by = .(ID = sourceid)]

str(uber_source_113)
```

Robin Lovelace, Jakub Nowosad, and Jannes Muenchow, [Geocomputation with R. Chapter 3: Attribute data operations](https://bookdown.org/robinlovelace/geocompr/attr.html)

```{r}
uber_source_113_nbds <- left_join(uber_nbds, uber_source_113, by = "ID")

str(uber_source_113_nbds)
```

```{r}
ggplot(uber_source_113_nbds) +
   geom_sf(aes(fill = MEAN_TRIP_DURATION), color = "white") +
  labs(title = "Mean Travel Time of Trips to Bay Street Corridor, 2018 Q2") +
  scale_fill_viridis(option="magma", name="Seconds", direction=-1)
```

```{r}
plot(uber_nbds[113,1])
```

## Playter Estates-Danforth

Playter Estates-Danforth is City of Toronto Neighbourhood 67 and Uber Movement Neighbourhood
122.

```{r}
uber_source_122 <- uber_movement[dstid == 122, .(MEAN_TRIP_DURATION = mean(mean_travel_time)), by = .(ID = sourceid)]

uber_source_122_nbds <- left_join(uber_nbds, uber_source_122, by = "ID")
```

```{r}
ggplot(uber_source_122_nbds) +
   geom_sf(aes(fill = MEAN_TRIP_DURATION), color = "white") +
  labs(title = "Mean Travel Time of Trips to Playter Estates-Danforth, 2018 Q2") +
  scale_fill_viridis(option="magma", name="Seconds", direction=-1)
```

```{r}
plot(uber_nbds[122,1])
```


# TTS 2016

[Data Management Group, University of Toronto Transportation Research Institute](http://dmg.utoronto.ca/open-data/download-tts-data)

**2016 TTS, City of Toronto**: <http://www.dmg.utoronto.ca/csv/tts/tts2016_ward_Toronto.csv>

Government of Ontario, [Transportation Tomorrow historical survey data](https://www.ontario.ca/data/transportation-tomorrow-historical-survey-data)

tts_dataset.zip

1986 TTS, 1991 TTS, 1996 TTS, 2001 TTS, 2006 TTS, 2011 TTS,
2016 TTS

## Data Manipulation and Variable Selection

```{r}
TTS2016_csv <- fread("tts2016_ward_Toronto.csv")

str(TTS2016_csv)
```

> * [55] "Total Number of persons"
> * [79] "Number of persons without a driver's licence"                                                 
> * [80] "Number of persons with a driver's licence"
> * [84] "Number of persons with a TTC metro pass"                                                      
> * [85] "Number of persons without any transit pass"   
> * [147] "Total number of trips made by residents in 24-hour"
> * [156] "Number of trips by residents by walk mode in 24-hour"
> * [157] "Number of trips by residents by bicycle in 24-hour"    
> * [416] "Number of transit trips made by residents with TTC Subway/RT"                                
> * [417] "Number of transit trips made by residents with TTC Bus/Streetcar" 

```{r}
TTS2016_selected <- TTS2016_csv[c(55,79,80,84,85,147,156,157,416,417),]

TTS2016_transpose <- transpose(TTS2016_selected)

v <- as.vector(as.matrix(TTS2016_transpose)[1,])

str(v)
```

```{r}
TTS2016_table <- TTS2016_transpose[-c(1:2),]

colnames(TTS2016_table) <- v
  
TTS2016_df <- as.data.frame(TTS2016_table)

for (y in v){
  TTS2016_df[,y] <- as.numeric(TTS2016_df[, y])
}

str(TTS2016_df)
```

```{r}
TTS2016_df$SCODE_NAME <- c(1:44)

str(TTS2016_df)
```

```{r}
TTS2016_df$TRIPS_PER_DAY <- TTS2016_df$`Total number of trips made by residents in 24-hour`/
  TTS2016_df$`Total Number of persons`

TTS2016_df$DRIVERS_LICENCE <- TTS2016_df$`Number of persons with a driver's licence`/
  TTS2016_df$`Total Number of persons`

TTS2016_df$TTC_PASS <- TTS2016_df$`Number of persons with a TTC metro pass`/
  TTS2016_df$`Total Number of persons`

TTS2016_df$WALKS_PER_DAY <- TTS2016_df$`Number of trips by residents by walk mode in 24-hour`/
  TTS2016_df$`Total Number of persons`

TTS2016_df$BIKES_PER_DAY <- TTS2016_df$`Number of trips by residents by bicycle in 24-hour`/
  TTS2016_df$`Total Number of persons`
```

## Plotting by Wards

[City of Toronto Wards](https://www.toronto.ca/city-government/data-research-maps/open-data/open-data-catalogue/#29b6fadf-0bd6-2af9-4a8c-8c41da285ad7)

```{r}
wards <- st_read("C:/Users/14165/Desktop/Shapefiles/May2010_WGS84/icitw_wgs84.shp", stringsAsFactors=FALSE)

str(wards)
```

```{r}
wards$SCODE_NAME <- as.integer(wards$SCODE_NAME)
```


```{r}
TTS2016_wards <- left_join(wards, TTS2016_df, by = "SCODE_NAME")

str(TTS2016_wards)
```

```{r}
ggplot(TTS2016_wards) +
   geom_sf(aes(fill = `Total Number of persons`), color = "white") +
  labs(title = "Total Number of persons by Ward, Toronto 2016")
```

```{r}
ggplot(TTS2016_wards) +
   geom_sf(aes(fill = TRIPS_PER_DAY), color = "white") +
  labs(title = "Average daily trips per residents by Ward, Toronto 2016")
```

```{r}
ggplot(TTS2016_wards) +
   geom_sf(aes(fill = DRIVERS_LICENCE), color = "white") +
  labs(title = "Ratio of persons with a driver's licence by Ward, Toronto 2016")
```

```{r}
ggplot(TTS2016_wards) +
   geom_sf(aes(fill = TTC_PASS), color = "white") +
  labs(title = "Ratio of persons with a TTC metro pass by Ward, Toronto 2016")
```

```{r}
ggplot(TTS2016_wards) +
   geom_sf(aes(fill = WALKS_PER_DAY), color = "white") +
  labs(title = "Average daily trips by walking per residents by Ward, Toronto 2016")
```

```{r}
ggplot(TTS2016_wards) +
   geom_sf(aes(fill = BIKES_PER_DAY), color = "white") +
  labs(title = "Average daily trips by bicycle per residents by Ward, Toronto 2016")
```

# Wellbeing Toronto

[Wellbeing Toronto - Transportation](https://www.toronto.ca/city-government/data-research-maps/open-data/open-data-catalogue/#cfa93ad9-7397-e851-f13e-f5097ed0fd55)

WB-Transportation.xlsx

```{r}
nbds <- st_read("C:/Users/14165/Desktop/Shapefiles/neighbourhoods_planning_areas_wgs84/NEIGHBORHOODS_WGS84.shp", stringsAsFactors=FALSE)
```

```{r}
nbds$AREA_S_CD <- as.integer(nbds$AREA_S_CD)
```

```{r}
WB <- read.csv("WB-Transportation.csv")

str(WB)
```

```{r}
WB_nbds <- left_join(nbds, WB, by = "AREA_S_CD")

str(WB_nbds)
```

## Plotting by Neighbourhoods

```{r}
ggplot(WB_nbds) +
   geom_sf(aes(fill = Traffic_Collisions), color = "white") +
  labs(title = "Traffic Collisions by Ward, Toronto 2011") +
  scale_fill_viridis(option="magma", name="Collisions", direction=-1) 
```

```{r}
ggplot(WB_nbds) +
   geom_sf(aes(fill = Non_Traffic_Collisions), color = "white") +
  labs(title = "Vehicle Collisions with Pedestrians/Other by Ward, Toronto 2011") +
    scale_fill_viridis(option="magma", name="Collisions", direction=-1) 

```


```{r}
ggplot(WB_nbds) +
   geom_sf(aes(fill = Daily_Road_Volume), color = "white") +
  labs(title = "Daily Road Volume by Ward, Toronto 2011") +
    scale_fill_viridis(option="magma", name="Road Volume", direction=-1) 
```