\documentclass{article}
\usepackage{amsmath,amssymb,mathrsfs,amsthm,xfrac}
\usepackage[LGR,T1]{fontenc}
\newcommand{\textgreek}[1]{\begingroup\fontencoding{LGR}\selectfont#1\endgroup}
%\usepackage{tikz-cd}
%\usepackage{hyperref}
\newcommand{\inner}[2]{\left\langle #1, #2 \right\rangle}
\newcommand{\tr}{\ensuremath\mathrm{tr}\,} 
\newcommand{\Span}{\ensuremath\mathrm{span}} 
\def\Re{\ensuremath{\mathrm{Re}}\,}
\def\Im{\ensuremath{\mathrm{Im}}\,}
\newcommand{\id}{\ensuremath\mathrm{id}} 
\newcommand{\gcm}{\ensuremath\mathrm{gcm}} 
\newcommand{\diam}{\ensuremath\mathrm{diam}} 
\newcommand{\sgn}{\ensuremath\mathrm{sgn}\,} 
\newcommand{\lcm}{\ensuremath\mathrm{lcm}} 
\newcommand{\supp}{\ensuremath\mathrm{supp}\,}
\newcommand{\dom}{\ensuremath\mathrm{dom}\,}
\newcommand{\norm}[1]{\left\Vert #1 \right\Vert}
\newcommand*\rfrac[2]{{}^{#1}\!/_{#2}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\begin{document}
\title{Extant medical writings}
\author{Jordan Bell\\ \texttt{jordan.bell@gmail.com}\\Department of Mathematics, University of Toronto}
\date{\today}

\maketitle

``On Theriac to Piso,'' Attributed to Galen: A Critical Edition with Translation and Commentary (Studies in Ancient Medicine). 
Robert Leigh

The Alexandrian Summaries of Galen's ``On Critical Days''. Gerrit Bos, Y. Tzvi Langermann

Galen, De diebus decretoriis, from Greek into Arabic: A Critical Edition, with Translation and Commentary, of Hunayn ibn Ishaq, Kitab ayyam al-buhran (Medicine in the Medieval Mediterranean). Glen M. Cooper

Galen, On Semen (Corpus Medicorum Graecorum). Philipp DeLacy

On the Therapeutic Method: Books I and II (Clarendon Later Ancient Philosophers). R. J. Hankinson

Galen on Bloodletting: A Study of the Origins, Development and Validity of his Opinions, with a Translation of the Three Works.
Peter Brain

Galen: On the Properties of Foodstuffs, Owen Powell

Galen: Method of Medicine, Loeb, Ian Johnston

Galen: On the Constitution of the Art of Medicine. The Art of Medicine. A Method of Medicine to Glaucon (Loeb Classical Library), Ian
Johnston

Galen: On the Natural Faculties (Loeb Classical Library), A. J. Brock

Galen, Selected Works, Oxford World Classics, Peter Singer

Galen: Psychological Writings: Avoiding Distress, Character Traits, The Diagnosis and Treatment of the Affections and Errors Peculiar to Each Person's ... of the Body (Cambridge Galen Translations). P. N. Singer

Galen: On Problematical Movements (Cambridge Classical Texts and Commentaries). Vivian Nutton

Galen: On Diseases and Symptoms. Ian Johnston

\end{document}